---
title: "Data 606 Final Project Proposal"
author: "Ken Popkin"
date: "3/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
#libraries
library(dplyr)
library(janitor)
```

### Data Preparation

```{r}
# load education data
education <- read.csv('https://raw.githubusercontent.com/krpopkin/Data606_Final_Project_Repository/master/01Data/Education(Graduation_Rates).csv')

cat('Dimension of raw data', dim(education),"\n")

#identify columns of interest and rename
education <- clean_names(education)
education <- select(education, 
                    state, 
                    county = area_name,
                    dropout = less_than_a_high_school_diploma_2014_18,
                    hs_diploma = high_school_diploma_only_2014_18,
                    some_college = some_college_or_associate_s_degree_2014_18,
                    four_year_degree = bachelor_s_degree_or_higher_2014_18,
                    dropout_percent = percent_of_adults_with_less_than_a_high_school_diploma_2014_18,
                    hs_diploma_percent = percent_of_adults_with_a_high_school_diploma_only_2014_18,
                    some_college_percent = percent_of_adults_completing_some_college_or_associate_s_degree_2014_18,
                    four_year_degree_percent = percent_of_adults_with_a_bachelor_s_degree_or_higher_2014_18)

cat('Dimension of selected data', dim(education), "\n")

#remove rows that are missing education data
education <- subset(education, dropout != '')
cat('Dimension of data after nulls removed', dim(education))
```
```{r}
head(education,1)

write.csv(education, 'education_clean.csv', row.names = FALSE)
```

```{r}
# load wages data
wages <- read.csv('https://raw.githubusercontent.com/krpopkin/Data606_Final_Project_Repository/master/01Data/allhlcn18.csv')

cat('Dimension of raw wage data:', dim(wages), '\n')

#filter to columns and rows of interest
#wages <- wages %>%
#  select(Area.Code, )
head(wages)

#write.csv(wages, 'wages_clean.csv', row.names = FALSE)
```

```{r}
#load home values data
homes <- read.csv('https://raw.githubusercontent.com/krpopkin/Data606_Final_Project_Repository/master/01Data/County_Zhvi_AllHomes(Home_Values).csv')

cat('Dimension of homes data:', dim(homes), '\n')

homes <- clean_names(homes)

write.csv(homes, 'homes_clean.csv', row.names = FALSE)

```


### Research question 
<I> Can education and wages be used as predictors for home values? </I>

There are many factors theorized to influence a home’s value and for this project we’ll attempt to answer if two popular topics, education and wages, can be used as predictors for a third popular topic, home values.

### Cases 

**What are the cases, and how many are there?**



### Data source and method of collection:
Data is collected from several sources listed below:

<B> Wages: </B> [Quarterly Census of Employment and Wages](https://www.bls.gov/cew/downloadable-data-files.htm)  
The Quarterly Census of Employment and Wages (QCEW) program publishes a quarterly count of employment and wages reported by employers covering more than 95 percent of U.S. jobs, available at the county, MSA, state and national levels by industry.

<B> Education: </B> [Local Education Agency](https://nces.ed.gov/ccd/f33agency.asp)  
Revenues and expenditures are audited after the close of the fiscal year and are then submitted to NCES by each state education agency.

<B> Home Values: </B> [Zillow](https://www.zillow.com/research/data/)  
Zillow Home Value Index (ZHVI) is a smoothed, seasonally adjusted measure of the typical home value and market changes across a given region and housing type. Zillow publishes top-tier ZHVI (typical value for homes within the 65th to 95th percentile range for a given region) and bottom-tier ZHVI (typical value for homes that fall within the 5th to 35th percentile range for a given region). 

### Type of study 
This is an observational study.

### Dependent Variable
The response variable is home values and is quantitative.

### Independent Variable
There are two independent variables; wages which is quantitative and education level which is qualitative.

### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}

```

