---
title: "DATA 606 Project Proposal (Education, Wages, and Home Values"
author: "Ken Popkin"
---

### Data Preparation

```{r setup, echo=TRUE, results='hide', warning=FALSE, message=FALSE}
# load data

```


### Research question 
<I> Can education and wages be used as predictors for home values? </I>

There are many factors theorized to influence a home’s value and for this project we’ll attempt to answer if two popular topics, education and wages, can be used as predictors for a third popular topic, home values.

### Cases 

**What are the cases, and how many are there?**



### Data source and method of collection:
Data is collected from several sources listed below:

<B> Wages: </B> [Quarterly Census of Employment and Wages](https://www.bls.gov/cew/downloadable-data-files.htm)  
The Quarterly Census of Employment and Wages (QCEW) program publishes a quarterly count of employment and wages reported by employers covering more than 95 percent of U.S. jobs, available at the county, MSA, state and national levels by industry.

<B> Education: </B> [Local Education Agency](https://nces.ed.gov/ccd/f33agency.asp)  
Revenues and expenditures are audited after the close of the fiscal year and are then submitted to NCES by each state education agency.

<B> Home Values: </B> [Zillow](https://www.zillow.com/research/data/)  
Zillow Home Value Index (ZHVI) is a smoothed, seasonally adjusted measure of the typical home value and market changes across a given region and housing type. Zillow publishes top-tier ZHVI (typical value for homes within the 65th to 95th percentile range for a given region) and bottom-tier ZHVI (typical value for homes that fall within the 5th to 35th percentile range for a given region). 

### Type of study 
This is an observational study.

### Dependent Variable
The response variable is home values and is quantitative.

### Independent Variable
There are two independent variables; wages which is quantitative and education level which is qualitative.

### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}

```

