---
title: "Data 606 Final Project Proposal"
author: "Ken Popkin"
date: "3/26/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
```

```{r}
#libraries
library(dplyr)
library(janitor)
library(sqldf)
library(stringr)
```

### Data Preparation

In this section we'll load education, wage, and housing data. For each load the features of interest will be selected, rows filtered as needed, and the results written to a clean_education, clean_wages, and clean_homes csv files.  There will also be a fourth csv file that merges all three files using "county" as the key field.

<I> Education data </I>
```{r}
# load education data
education <- read.csv('https://raw.githubusercontent.com/krpopkin/Data606_Final_Project_Repository/master/01Data/Education(Graduation_Rates).csv')

cat('Dimension of raw education data', dim(education),"\n")

#identify columns of interest and rename
education <- clean_names(education)
education <- select(education, 
                    state, 
                    county = area_name,
                    dropout = less_than_a_high_school_diploma_2014_18,
                    hs_diploma = high_school_diploma_only_2014_18,
                    some_college = some_college_or_associate_s_degree_2014_18,
                    four_year_degree = bachelor_s_degree_or_higher_2014_18,
                    dropout_percent = percent_of_adults_with_less_than_a_high_school_diploma_2014_18,
                    hs_diploma_percent = percent_of_adults_with_a_high_school_diploma_only_2014_18,
                    some_college_percent = percent_of_adults_completing_some_college_or_associate_s_degree_2014_18,
                    four_year_degree_percent = percent_of_adults_with_a_bachelor_s_degree_or_higher_2014_18)

cat('Dimension of selected education data', dim(education), "\n")

#remove rows that are missing education data
education <- subset(education, dropout != '')

#add the full state name to the education dataframe
state_mapping <- read.csv('https://raw.githubusercontent.com/krpopkin/Data606_Final_Project_Repository/master/01Data/state_mapping.csv')

education <- sqldf('SELECT * 
                    FROM education
                    JOIN state_mapping ON education.state = state_mapping.state_short')

cat('Dimension of education data after cleaning and filtering', dim(education))
```
```{r}
head(education,1)

path_out = 'C:\\Users\\user\\Documents\\00_Applications_DataScience\\CUNY\\DATA606\\Data606_Final_Project_Repository\\01Data\\'
write.csv(education, paste(path_out, 'clean_education.csv'), row.names = FALSE)
```

<I> Wages data </I>
```{r}
# load wages data
wages <- read.csv('https://raw.githubusercontent.com/krpopkin/Data606_Final_Project_Repository/master/01Data/allhlcn18.csv')

cat('Dimension of raw wage data:', dim(wages), '\n')

#filter to columns and rows of interest
wages <- wages %>%
  select(type = Area.Type,
         state = St.Name, 
         county = Area, 
         ownership = Ownership,
         resident_count = Annual.Average.Establishment.Count,
         avg_annual_pay = Annual.Average.Pay) %>%
  subset(type == 'County' & ownership == 'Total Covered') %>%
  select(-type)

#Remove the state name from the data in the county column,ie "Bibb County, Alabama" becomes "Bibb County"
wages$position <- (regexpr(pattern =',', wages$county)) - 1
wages$county <- str_sub(wages$county, end = wages$position)

cat('Dimension of data after filtering', dim(wages))
```

```{r}
head(wages,1)
write.csv(wages, paste(path_out, 'clean_wages.csv'), row.names = FALSE)
```

<I> Home values data </I>
```{r}
#load home values data
homes <- read.csv('https://raw.githubusercontent.com/krpopkin/Data606_Final_Project_Repository/master/01Data/County_Zhvi_AllHomes(Home_Values).csv')

homes <- clean_names(homes)

cat('Dimension of homes data:', dim(homes), '\n')

homes <-  homes %>%
  select(county = region_name, state = state, metro = metro,
         jan = x2018_01, feb = x2018_02, mar = x2018_03, apr = x2018_04, may = x2018_05, jun = x2018_06,
         jul = x2018_07, aug = x2018_08, sep = x2018_09, oct = x2018_10, nov = x2018_11, dec = x2018_12)

homes$avg_home_value = (homes$jan + homes$feb + homes$mar + homes$apr + homes$may + homes$jun + homes$jul + homes$aug + homes$sep + homes$oct + homes$nov + homes$dec)/12

cat('Dimension of selected homes data:', dim(homes))
```

```{r}
head(homes,1)
write.csv(homes, paste(path_out, 'clean_homes.csv'), row.names = FALSE)
```

<I> Join files </I>
Now we'll join the education, wages, and housing dataframes using state and county as the keys.  This will give us a single data file to complete the project with.
```{r}
#Join education and wages
masterdata <- sqldf('SELECT * 
                    FROM education
                    INNER JOIN wages ON (education.county = wages.county AND education.state_long = wages.state)')

#Join masterdata to homes
masterdata <- sqldf('SELECT *
                     FROM masterdata
                     INNER JOIN homes ON )

dim(masterdata)
write.csv(masterdata, paste(path_out, 'master.csv'), row.names = FALSE)
```


### Research question 
<I> Can education and wages be used as predictors for home values? </I>

There are many factors theorized to influence a home’s value and for this project we’ll attempt to answer if two popular topics, education and wages, can be used as predictors for a third popular topic, home values.

### Cases 

**What are the cases, and how many are there?**



### Data source and method of collection:
Data is collected from several sources listed below:

<B> Wages: </B> [Quarterly Census of Employment and Wages](https://www.bls.gov/cew/downloadable-data-files.htm)  
The Quarterly Census of Employment and Wages (QCEW) program publishes a quarterly count of employment and wages reported by employers covering more than 95 percent of U.S. jobs, available at the county, MSA, state and national levels by industry.

<B> Education: </B> [Local Education Agency](https://nces.ed.gov/ccd/f33agency.asp)  
Revenues and expenditures are audited after the close of the fiscal year and are then submitted to NCES by each state education agency.

<B> Home Values: </B> [Zillow](https://www.zillow.com/research/data/)  
Zillow Home Value Index (ZHVI) is a smoothed, seasonally adjusted measure of the typical home value and market changes across a given region and housing type. Zillow publishes top-tier ZHVI (typical value for homes within the 65th to 95th percentile range for a given region) and bottom-tier ZHVI (typical value for homes that fall within the 5th to 35th percentile range for a given region). 

### Type of study 
This is an observational study.

### Dependent Variable
The response variable is home values and is quantitative.

### Independent Variable
There are two independent variables; wages which is quantitative and education level which is qualitative.

### Relevant summary statistics 

**Provide summary statistics for each the variables. Also include appropriate visualizations related to your research question (e.g. scatter plot, boxplots, etc). This step requires the use of R, hence a code chunk is provided below. Insert more code chunks as needed.**

```{r}

```

